---
title: "Retail Demand Forecasting — R Analysis"
author: "Demand Forecasting for Retail Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

# 1. Purpose

This report provides an R-based analytical and validation view of the retail
demand forecasting project.

It consumes the compact analytical outputs produced by Phase 13 and the stored
Phase 12 validation forecasts rather than reloading the full raw retail
dataset. Every figure below is produced by the audited workflow in
`r/r_analysis.R`; nothing is entered by hand.

Python remains the primary forecasting implementation. R is used here as an
independent analytical, validation and reporting workflow.

# 2. Method

The report sources the phase script and runs it. That means the report and the
script verify the same evidence, and the report cannot render if any quality
check fails.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(yardstick)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

find_project_root <- function(start_dir = getwd()) {
  current <- normalizePath(start_dir, winslash = "/", mustWork = TRUE)

  repeat {
    if (
      file.exists(file.path(current, "pyproject.toml")) &&
        dir.exists(file.path(current, "docs"))
    ) {
      return(current)
    }

    parent <- dirname(current)

    if (identical(parent, current)) {
      stop("Could not locate the project root from the knitting directory.")
    }

    current <- parent
  }
}

project_root <- find_project_root()

# Pandoc resolves resources relative to this document's directory, so figures
# are referenced relative to the report rather than by absolute path.
relative_to_report <- function(path) {
  input_dir <- normalizePath(
    dirname(knitr::current_input(dir = TRUE)),
    winslash = "/",
    mustWork = TRUE
  )

  root_prefix <- paste0("^", project_root, "/?")
  depth <- length(strsplit(
    sub(root_prefix, "", input_dir),
    "/"
  )[[1]])

  file.path(
    paste(rep("..", depth), collapse = "/"),
    sub(root_prefix, "", normalizePath(path, winslash = "/", mustWork = FALSE))
  )
}

source(
  file.path(project_root, "r", "r_analysis.R")
)

results <- main()
```

# 3. Validation Result

The workflow records every validation in a machine-readable quality report and
stops if any check fails.

```{r quality-summary}
quality <- results$quality

summary_table <- tibble(
  checks = nrow(quality),
  passed = sum(quality$passed),
  failed = sum(!quality$passed)
)

summary_table
```

```{r quality-failures}
quality %>%
  filter(!passed) %>%
  select(check, actual, expected)
```

# 4. Input Coverage

```{r input-coverage}
coverage <- tibble(
  input = c(
    "Historical demand summary",
    "Demand variability summary",
    "Historical inventory scenarios",
    "Forecast-error inventory scenarios",
    "Forecast-error summary",
    "Densification summary",
    "Phase 13 insights",
    "Phase 12 tuned validation results",
    "Phase 12 stored predictions"
  ),
  rows = c(
    nrow(results$inputs$demand),
    nrow(results$inputs$variability),
    nrow(results$scenarios),
    nrow(results$forecast_scenarios),
    nrow(results$inputs$forecast_errors),
    nrow(results$inputs$densification),
    nrow(results$inputs$insights),
    nrow(results$inputs$validation_results),
    nrow(results$inputs$predictions)
  )
)

coverage
```

# 5. Store-Level Demand

The analysis uses `dplyr` to join the demand and variability summaries and to
arrange the stores by average demand.

```{r store-analysis}
store_analysis <- results$store_analysis

store_analysis %>%
  select(
    store_id,
    training_days,
    mean_daily_demand,
    median_daily_demand,
    min_daily_demand,
    max_daily_demand,
    std_daily_demand,
    coefficient_of_variation
  )
```

## Highest Average Demand

```{r highest-demand}
highest_demand <- store_analysis %>%
  slice_max(
    order_by = mean_daily_demand,
    n = 1,
    with_ties = FALSE
  )

highest_demand
```

The highest observed average daily demand occurs at Store
`r highest_demand$store_id`, at
`r format(round(highest_demand$mean_daily_demand, 3), big.mark = ",")` units
per day over `r highest_demand$training_days` densified training days.

## Highest Relative Variability

```{r highest-variability}
highest_variability <- store_analysis %>%
  slice_max(
    order_by = coefficient_of_variation,
    n = 1,
    with_ties = FALSE
  )

highest_variability
```

Store `r highest_variability$store_id` has the highest observed coefficient of
variation (`r round(highest_variability$coefficient_of_variation, 6)`). Its
history is the shortest of the four stores, so this estimate rests on fewer
observations than the others.

## Visualization

The following chart uses a `ggplot2` layered construction with a data mapping
and geometric objects.

```{r demand-plot}
ggplot(
  store_analysis,
  aes(
    x = factor(store_id),
    y = mean_daily_demand
  )
) +
  geom_col() +
  labs(
    title = "Average Daily Demand by Store",
    subtitle = "Densified training store-day series (Phases 11-13 basis)",
    x = "Store",
    y = "Mean Daily Demand"
  ) +
  theme_minimal()
```

```{r variability-plot}
ggplot(
  store_analysis,
  aes(
    x = factor(store_id),
    y = coefficient_of_variation
  )
) +
  geom_col() +
  labs(
    title = "Relative Demand Variability by Store",
    subtitle = "Coefficient of variation (standard deviation / mean)",
    x = "Store",
    y = "Coefficient of Variation"
  ) +
  theme_minimal()
```

# 6. Exported Figures

The workflow exports the same evidence with `ggsave()`.

```{r exported-figures, out.width="100%"}
plot_files <- c(
  "average_daily_demand_by_store.png",
  "demand_variability_by_store.png",
  "inventory_reorder_point_scenarios.png",
  "inventory_scenario_family_comparison.png"
)

knitr::include_graphics(
  vapply(
    file.path(results$paths$plot_dir, plot_files),
    relative_to_report,
    character(1)
  )
)
```

# 7. Inventory Scenario Analysis

The inventory calculations are planning scenarios rather than operational
requirements because supplier lead times and actual service-level targets were
not available.

```{r scenario-summary}
results$scenario_summary
```

```{r forecast-scenario-summary}
results$forecast_scenario_summary
```

## 14-Day / 95% Planning Scenario

```{r baseline-scenarios}
results$historical_baseline %>%
  select(
    store_id,
    mean_daily_demand,
    expected_lead_time_demand,
    safety_stock,
    reorder_point
  )

results$forecast_baseline %>%
  select(
    store_id,
    bias_adjusted_daily_demand,
    expected_lead_time_demand,
    safety_stock,
    reorder_point
  )
```

## Faceted Scenario Comparison

`facet_wrap()` provides a separate view per store and separates the two
uncertainty inputs.

```{r scenario-facets}
bind_rows(
  results$scenarios %>%
    mutate(family = "historical variability"),
  results$forecast_scenarios %>%
    mutate(family = "forecast residual error")
) %>%
  ggplot(
    aes(
      x = lead_time_days,
      y = reorder_point,
      colour = family,
      linetype = factor(service_level)
    )
  ) +
  geom_line() +
  geom_point() +
  facet_wrap(~ store_id) +
  labs(
    title = "Reorder-Point Scenarios by Lead Time and Service Level",
    x = "Lead Time (Days)",
    y = "Reorder Point",
    colour = "Uncertainty input",
    linetype = "Service Level"
  ) +
  theme_minimal()
```

# 8. Phase 13 Reconciliation

R independently reproduces every insight Phase 13 published, and reconciles
the published value as well as the store label.

```{r phase13-reconciliation}
results$reconciliation %>%
  select(
    insight_type,
    store_id,
    metric,
    phase13_value,
    r_value,
    relative_difference,
    match
  )
```

```{r phase13-summary}
tibble(
  rows_reconciled = nrow(results$reconciliation),
  rows_matching = sum(results$reconciliation$match),
  maximum_relative_difference = max(
    results$reconciliation$relative_difference
  )
)
```

# 9. Phase 12 Forecast-Verification Reconciliation

The workflow recomputes the validation metrics from the 456 stored Phase 12
validation forecasts using `yardstick`, the tidymodels metrics package, and
reconciles them with the reported values.

```{r metric-reconciliation}
results$metrics %>%
  select(
    store_id,
    observations,
    rmse,
    mae,
    mape_percent,
    reported_rmse,
    reported_mape_percent,
    rmse_relative_difference,
    mape_relative_difference
  )
```

Mean absolute error is not published by Phase 12, so it is R-derived evidence
rather than a reconciliation.

# 10. Densification Reconciliation

```{r densification}
results$densification
```

The densified training series used by Phases 11-13 is reproduced exactly:
`r sum(results$densification$densified_days)` store-days, of which
`r sum(results$densification$synthetic_days)` is a single zero-filled day.

# 11. Environment

```{r environment}
results$environment
```

```{r session-info}
sessionInfo()
```

# 12. Interpretation

The R analysis independently reproduces the principal Phase 13 store-level
findings and the Phase 12 validation metrics:

- Store `r highest_demand$store_id` has the highest observed average daily
  physical demand.
- Store `r highest_variability$store_id` has the highest relative demand
  variability, on the shortest history.
- Every published Phase 13 insight reconciles with the independently
  recomputed value.
- Recomputed RMSE and MAPE agree with the Phase 12 artifacts, so the forecast
  evidence is verified in a second language.
- Inventory scenarios vary with both assumed lead time and service level, and
  the forecast-error family sizes buffers from realised forecast error rather
  than raw historical spread.
- These inventory values should not be treated as actual procurement
  requirements without operational information.

# 13. Limitations

The R analysis does not create a competing production forecasting model.

The validated forecasting models remain those selected through the Python
workflow in Phases 11 and 12.

The dataset does not provide:

- supplier lead times;
- operational service-level requirements;
- holding costs;
- ordering costs;
- stockout costs;
- current inventory;
- purchase orders;
- warehouse capacity.

Therefore, inventory values are scenario estimates.

Store 4's history and its single cross-validation fold mean its variability and
forecast-error estimates rest on fewer observations than the other stores, and
that caveat is carried in the generated artifacts rather than only in prose.

# 14. Reproducibility

The report is generated from project files rather than manually entered
results.

```{r quality-gate}
if (any(!results$quality$passed)) {
  stop("Phase 14 quality checks failed; the report must not render.")
}

cat(
  "All",
  nrow(results$quality),
  "quality checks passed before this report was rendered.\n"
)
```

The report can be regenerated from the repository root with:

```text
Rscript r/r_analysis.R
Rscript -e "rmarkdown::render('r/r_analysis_report.Rmd', output_dir = file.path(getwd(), 'data', 'analysis', 'r'))"
```
