---
title: "Retail Demand Forecasting — R Analysis"
author: "Demand Forecasting for Retail Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

# 1. Purpose

This report provides an R-based analytical view of the retail demand
forecasting project.

The analysis uses the compact analytical outputs produced during Phase 13
rather than reloading the full raw retail dataset.

Python remains the primary forecasting implementation. R is used here to
demonstrate an independent analytical and reporting workflow.

# 2. Packages

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidymodels)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

# 3. Load Phase 13 Results

```{r load-data}
analysis_dir <- file.path(
  "..",
  "data",
  "analysis"
)

demand <- read_csv(
  file.path(
    analysis_dir,
    "inventory_demand_summary.csv"
  ),
  show_col_types = FALSE
)

variability <- read_csv(
  file.path(
    analysis_dir,
    "inventory_variability_summary.csv"
  ),
  show_col_types = FALSE
)

scenarios <- read_csv(
  file.path(
    analysis_dir,
    "inventory_scenarios.csv"
  ),
  show_col_types = FALSE
)
```

# 4. Store-Level Demand

The following analysis uses `dplyr` to combine the demand and variability
summaries.

```{r store-analysis}
store_analysis <- demand %>%
  left_join(
    variability %>%
      select(
        store_id,
        std_daily_demand,
        p90_daily_demand,
        p95_daily_demand,
        p99_daily_demand
      ),
    by = "store_id"
  ) %>%
  arrange(desc(mean_daily_demand))

store_analysis
```

## Highest Average Demand

```{r highest-demand}
highest_demand <- store_analysis %>%
  slice_max(
    mean_daily_demand,
    n = 1,
    with_ties = FALSE
  )

highest_demand
```

The highest observed average daily demand occurs at Store
`r highest_demand$store_id`.

# 5. Demand Variability

```{r highest-variability}
highest_variability <- store_analysis %>%
  slice_max(
    coefficient_of_variation,
    n = 1,
    with_ties = FALSE
  )

highest_variability
```

Store `r highest_variability$store_id` has the highest observed coefficient
of variation.

## Visualization

The following chart uses a `ggplot2` layered construction with a data
mapping and geometric object.

```{r demand-plot}
ggplot(
  store_analysis,
  aes(
    x = factor(store_id),
    y = mean_daily_demand
  )
) +
  geom_col() +
  labs(
    title = "Average Daily Demand by Store",
    x = "Store",
    y = "Mean Daily Demand"
  ) +
  theme_minimal()
```

# 6. Relative Variability

```{r variability-plot}
ggplot(
  store_analysis,
  aes(
    x = factor(store_id),
    y = coefficient_of_variation
  )
) +
  geom_col() +
  labs(
    title = "Relative Demand Variability by Store",
    x = "Store",
    y = "Coefficient of Variation"
  ) +
  theme_minimal()
```

# 7. Inventory Scenario Analysis

The inventory calculations are planning scenarios rather than operational
requirements because supplier lead times and actual service-level targets
were not available.

```{r scenario-summary}
scenario_summary <- scenarios %>%
  group_by(
    lead_time_days,
    service_level
  ) %>%
  summarise(
    stores = n_distinct(store_id),
    average_reorder_point = mean(reorder_point),
    maximum_reorder_point = max(reorder_point),
    average_safety_stock = mean(safety_stock),
    .groups = "drop"
  )

scenario_summary
```

## Reorder-Point Visualization

```{r reorder-point-plot}
ggplot(
  scenarios,
  aes(
    x = lead_time_days,
    y = reorder_point,
    group = service_level,
    linetype = factor(service_level)
  )
) +
  geom_line() +
  geom_point() +
  facet_wrap(~ store_id) +
  labs(
    title = "Scenario Reorder Points",
    x = "Lead Time (Days)",
    y = "Reorder Point",
    linetype = "Service Level"
  ) +
  theme_minimal()
```

The use of `facet_wrap()` provides separate views for each store.

# 8. Phase 13 Consistency Check

The R analysis should reproduce the principal store-level findings already
established in Phase 13.

```{r consistency}
phase13_insights <- read_csv(
  file.path(
    analysis_dir,
    "forecast_inventory_insights.csv"
  ),
  show_col_types = FALSE
)

phase13_highest_demand <- phase13_insights %>%
  filter(
    insight_type == "highest_average_demand"
  ) %>%
  slice(1)

phase13_highest_variability <- phase13_insights %>%
  filter(
    insight_type == "highest_relative_variability"
  ) %>%
  slice(1)

consistency <- tibble(
  check = c(
    "Highest average demand store",
    "Highest relative variability store"
  ),
  phase_13 = c(
    phase13_highest_demand$store_id,
    phase13_highest_variability$store_id
  ),
  r_analysis = c(
    highest_demand$store_id,
    highest_variability$store_id
  )
) %>%
  mutate(
    match = phase_13 == r_analysis
  )

consistency
```

# 9. Interpretation

The R analysis independently reproduces the principal Phase 13 store-level
findings.

The analysis shows that:

- Store `r highest_demand$store_id` has the highest observed average daily
  physical demand.
- Store `r highest_variability$store_id` has the highest relative demand
  variability.
- Inventory scenarios vary with both assumed lead time and service level.
- These inventory values should not be treated as actual procurement
  requirements without operational information.

# 10. Limitations

The R analysis does not create a competing production forecasting model.

The validated forecasting models remain those selected through the Python
workflow in Phases 11 and 12.

The dataset does not provide:

- supplier lead times;
- operational service-level requirements;
- holding costs;
- ordering costs;
- stockout costs;
- current inventory;
- purchase orders;
- warehouse capacity.

Therefore, inventory values are scenario estimates.

# 11. Reproducibility

The report is generated from project files rather than manually entered
results.

The R script and this R Markdown document are version-controlled project
artifacts.

The report can be regenerated by running the Phase 14 R workflow from the
repository root.